How the PyPy JIT works

### 2012-03-07 talk/pycon2012/jit/slides.tex
also this
hopefully make this clearer

\begin{frame}{Generating the JIT}
\begin{itemize}
\item low-level flowgraphs converted to jitcodes
\item interpreter decides what functions to ``look into''
\item jitcodes stored in binary
\end{itemize}
\end{frame}

@@

\begin{frame}[fragile]{JIT IR example - Head of loop}
\footnotesize{
\begin{verbatim}
# arguments to loop
[p0, p1, p2, p3, i4, p5, i6, i7, p8, p9, p10, p11, p12, p13]
#25 LOAD_FAST
guard_value(i6, 3)
guard_nonnull(p10)
guard_value(i4, 0)
#28 LOAD_FAST
guard_nonnull(p9)
#31 COMPARE_OP
guard_class(p10, W_IntObject)
guard_class(p9, W_IntObject)
i18 = getfield_gc_pure(p10, <W_IntObject.inst_intval>)
i19 = getfield_gc_pure(p9, <W_IntObject.inst_intval>)
i20 = int_le(i18, i19)
guard_true(i20)
#34 POP_JUMP_IF_FALSE
\end{verbatim}
}
\end{frame}

\begin{frame}[fragile]{JIT IR example - computing modulo}
\footnotesize{
\begin{verbatim}
#37 LOAD_FAST
guard_nonnull(p8)
#40 LOAD_FAST
#43 BINARY_MODULO
guard_class(p8, W_IntObject)
i22 = getfield_gc_pure(p8, <W_IntObject.inst_intval>)
i23 = int_is_zero(i18)
guard_false(i23)
i25 = int_eq(i22, -9223372036854775808)
guard_false(i25)
i26 = int_mod(i22, i18)
i28 = int_lt(i18, 0)
guard_false(i28)
i30 = int_rshift(i26, 63)
i31 = int_and(i18, i30)
i32 = int_add(i26, i31)
p34 = new_with_vtable(W_IntObject)
setfield_gc(p34, i32, <W_IntObject.inst_intval>)
\end{verbatim}
}
\end{frame}

\begin{frame}[fragile]{JIT IR example - check modulo}
\footnotesize{
\begin{verbatim}
#44 LOAD_CONST
guard_value(p3, ConstPtr(ptr35))
#47 COMPARE_OP')
i37 = int_eq(i32, 0)
guard_false(i37)
#50 POP_JUMP_IF_FALSE
\end{verbatim}
}
\end{frame}

\begin{frame}[fragile]{JIT IR example - finish loop}
\footnotesize{
\begin{verbatim}
#57 LOAD_FAST
#60 LOAD_FAST
#63 INPLACE_ADD
i38 = getfield_gc_pure(p10, <W_IntObject.inst_intval>)
i40 = int_add_ovf(i38, 1)
guard_no_overflow()
p42 = new_with_vtable(W_IntObject)
setfield_gc(p42, i40, <W_IntObject.inst_intval>)
#64 STORE_FAST
#67 JUMP_ABSOLUTE
\end{verbatim}
}
\end{frame}

\begin{frame}{Optimizations}
\begin{itemize}
\item limited by speed
\item classical compiler optimizations
\begin{itemize}
\item constant folding
\item strength reduction
\item integer bounds
\end{itemize}
\item most important is removing indirection: virtuals and virtualizables
\end{itemize}
\end{frame}

>>>>

\begin{frame}{Generating the JIT}
\begin{itemize}
\item RPython interpreter is in low-level flowgraphs
\item low-level flowgraphs converted to jitcodes
\item interpreter decides what functions to ``look into''
\item jitcodes stored in binary
\end{itemize}
\end{frame}

@@

\begin{frame}[fragile]{JIT IR example - Head of loop}
\footnotesize{
\begin{verbatim}
# arguments to loop
[p0, p1, p2, p3, i4, p5, i6, i7, p8, p9, p10, p11, p12, p13]
# while i < limit
#25 LOAD_FAST
guard_value(i6, 3)
guard_nonnull(p10)
guard_value(i4, 0)
#28 LOAD_FAST
guard_nonnull(p9)
#31 COMPARE_OP
guard_class(p10, W_IntObject)
guard_class(p9, W_IntObject)
i18 = getfield_gc_pure(p10, <W_IntObject.inst_intval>)
i19 = getfield_gc_pure(p9, <W_IntObject.inst_intval>)
i20 = int_le(i18, i19)
guard_true(i20)
#34 POP_JUMP_IF_FALSE
\end{verbatim}
}
\end{frame}

\begin{frame}[fragile]{JIT IR example - computing modulo}
\footnotesize{
\begin{verbatim}
# n % i
#37 LOAD_FAST
guard_nonnull(p8)
#40 LOAD_FAST
#43 BINARY_MODULO
guard_class(p8, W_IntObject)
i22 = ((W_IntObject)p8)->inst_intval)
i23 = int_is_zero(i18)
guard_false(i23)
i25 = int_eq(i22, -9223372036854775808)
guard_false(i25)
i26 = int_mod(i22, i18)
i28 = int_lt(i18, 0)
guard_false(i28)
i30 = int_rshift(i26, 63)
i31 = int_and(i18, i30)
i32 = int_add(i26, i31)
p34 = new_with_vtable(W_IntObject)
((W_IntObject)p34)->inst_intval = i32
\end{verbatim}
}
\end{frame}

\begin{frame}[fragile]{JIT IR example - check modulo}
\footnotesize{
\begin{verbatim}
# if <modulo> == 0:
#44 LOAD_CONST
guard_value(p3, ConstPtr(ptr35))
#47 COMPARE_OP')
i37 = int_eq(i32, 0)
guard_false(i37)
#50 POP_JUMP_IF_FALSE
\end{verbatim}
}
\end{frame}

\begin{frame}[fragile]{JIT IR example - finish loop}
\footnotesize{
\begin{verbatim}
# i += 1
#57 LOAD_FAST
#60 LOAD_FAST
#63 INPLACE_ADD
i38 = ((W_IntObject)p10)->inst_intval
i40 = int_add_ovf(i38, 1)
guard_no_overflow()
p42 = new_with_vtable(W_IntObject)
((W_IntObject)p42)->inst_inval = i40
#64 STORE_FAST
#67 JUMP_ABSOLUTE
\end{verbatim}
}
\end{frame}

\begin{frame}{Optimizations}
\begin{itemize}
\item limited by speed
\item classical compiler optimizations
\begin{itemize}
\item constant folding
\item strength reduction
\item integer bounds
\end{itemize}
\item remove extra gaurds
\item most important is removing indirection: virtuals and virtualizables
\end{itemize}
\end{frame}
